---
title: "index_throw_away_code"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>
<center>
```{r echo=FALSE, results="hide", include=FALSE, message=FALSE}
library(tigris)
library(leaflet)
library(plyr)
library(dplyr)
library(widgetframe)
library(RColorBrewer)
library(htmlwidgets)
library(htmltools)

load("/Users/carlosalgara/Dropbox/Github_Website/calgara.github.io/Rshinyapp/shiny_data/clean_shiny_data.Rdata")
```
```{r echo=FALSE, width = "800px", height = "400px", out.width = "80%"}
map3
```
</center>
<br>
Please visit my [Research Agenda Statement](Algara_Research_Statement.pdf) for an overview of my broad research agenda, current projects in this agenda, and future directions. To learn more about ongoing working projects of this research agenda, please visit the [Publications & Working Papers](https://calgara.github.io/working_papers.html) and [CV](https://calgara.github.io/cv.html) tabs of this website. To learn more about my teaching philosophy and comprehensive evidence of teaching effectiveness, please visit the [Teaching Porfolio](https://calgara.github.io/teaching.html).    